---
title: "Reproducible Research: Peer Assessment 1"
author: "EvenGu"
output: 
  html_document:
    keep_md: true
---

```{r setoptions}
## set global options to 1) supress package loading message and 2) format the numbers
library(knitr)
opts_chunk$set(message=FALSE, options(digits=2))
```

## Loading and preprocessing the data
```{r}
unzip(zipfile="activity.zip", overwrite=T)
dF <- read.csv("activity.csv",header=T)

library(ggplot2)
library(dplyr)

```


## What is mean total number of steps taken per day?
```{r}
dailySteps <- dF %>% group_by(date) %>%
  summarize(daily.steps=sum(steps,na.rm=TRUE))
meanTotal <- mean(dailySteps$daily.steps)
medianTotal <- median(dailySteps$daily.steps)
```
Here is the plot.
```{r plot1, fig.height=4}
qplot(daily.steps, data=dailySteps) + ggtitle("Total Number of Steps per Day")
```

**The mean total number of steps taken per day is `r meanTotal`.**
**The median is `r medianTotal`.**


## What is the average daily activity pattern?
```{r}
timeSteps <- dF %>% group_by(interval) %>%
  summarize(activity.pattern=mean(steps,na.rm=TRUE))
time <- sprintf("%04d",timeSteps$interval) 
timeSteps$time <- gsub('^(.{2})(.*)$', '\\1:\\2', time)
id <- which.max(timeSteps$activity.pattern)
```
Here is the plot.
```{r plot2, fig.height=4}
plot(as.factor(timeSteps$time), timeSteps$activity.pattern)
lines(as.factor(timeSteps$time), timeSteps$activity.pattern)
title(main="Average Daily Activity Pattern", xlab="Time", ylab="Average Steps")
```

**The time interval that contains the maximum number of steps, when averaged across all the days in the dataset, is from `r timeSteps$time[id-1]` to `r timeSteps$time[id]`.**


## Imputing missing values
```{r imputNA}
naTotal <- sum(is.na(dF$steps))

## imputing NAs using the mean for that 5-minute interval across all the days
l=split(dF,as.factor(dF$interval))
for (i in 1:length(l)) {
  l[[i]]$steps[is.na(l[[i]]$steps)] = timeSteps$activity.pattern[i]
}
dF2 = unsplit(l,as.factor(dF$interval))

dailySteps.imput <- dF2 %>% group_by(date) %>%
  summarize(daily.steps=sum(steps))
meanTotal.imput <- mean(dailySteps.imput$daily.steps)
medianTotal.imput <- median(dailySteps.imput$daily.steps)
```
Here is the plot.
```{r plot3, fig.height=4}
qplot(daily.steps, data=dailySteps.imput) +
  ggtitle("Total Number of Steps per Day with NAs imputed")
```

**The total number of missing values (NA) in the dataset is `r naTotal`.**
**After imputing missing data, the mean total number of steps taken per day is `r format(meanTotal.imput,scientific=FALSE)` and the median is `r format(medianTotal.imput,scientific=FALSE)`. They are different from the estimates of the first part of the assignment.**


## Are there differences in activity patterns between weekdays and weekends?
```{r}
dF2$days <- weekdays(as.POSIXct(dF2$date))
wDay <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
dF2$weekday <- factor(dF2$days %in% wDay, 
         levels=c(FALSE,TRUE), labels=c('weekend', 'weekday'))

timeSteps2 <- dF2 %>% group_by(weekday,interval) %>%
  summarize(activity.pattern=mean(steps,na.rm=TRUE))
```
Here is the plot.
```{r plot4, fig.height=4}
qplot(x=interval, y=activity.pattern, data=timeSteps2, geom="line", facets=weekday~.) + 
  ylab("Number of Averaged Steps")
```




